---
title: "Statistical Analysis on UCI Heart Disease Data - Probability & Stats Project"
author: "Abhijit - Saksham - Sharvin"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r full-analysis, echo=TRUE}

## R Markdown
setwd("D:/Sharvin  Jio Institute/Probability & Stats!/et/")

#Load the dataset
# The dataset is from kaggle. 
# Link : https://www.kaggle.com/datasets/redwankarimsony/heart-disease-data/

heart_disease=read.csv("heart_disease_uci.csv")
head(heart_disease)

numerical_cols = c("age", "trestbps", "chol", "thalch", "oldpeak", "ca")

#Mean of Age
mean_age=mean(heart_disease$age, na.rm = TRUE)
mean_age
summary(heart_disease)
# <!-- We have 8 categorical columns and 6 numerical columns. -->
# <!-- Most patients are from the Cleveland dataset. -->
# <!-- Typical angina and asymptomatic chest pain types are the most common. -->
# <!-- Most patients have 0 major vessels colored during fluoroscopy. -->
# <!-- The mean values are: -->
# <!-- Age: ~53.5 years -->
# <!-- Resting BP: ~132 -->
# <!-- Cholesterol: ~199 -->
# <!-- Max Heart Rate: ~137 -->
# <!-- ST Depression (Oldpeak): ~0.88 -->

#Finding for any Co-relation between the given dataset

numerical_cols = c("age", "trestbps", "chol", "thalch", "oldpeak", "ca")
num_data=heart_disease[,numerical_cols]
cor_matrix= cor(num_data,use="complete.obs")
print(cor_matrix)

#Correlation Analysis
# <!-- age is negatively correlated with thalch (maximum heart rate) → as age increases, heart rate decreases. -->
# <!-- age is positively correlated with ca (number of blocked vessels) and oldpeak (ST depression). -->
# <!-- trestbps is weakly correlated with oldpeak. -->
# <!-- chol is weakly correlated with thalch. -->
# <!-- ca is moderately correlated with oldpeak. -->

# 2. One Sample and Two Sample T-Test
t.test(heart_disease$chol, mu = 200)
# <!-- p is high → null will fly. -->
# <!-- Since p-value > 0.05, we accept the null hypothesis. -->
# <!-- There is no significant difference between the sample mean and the assumed value of 200 mg/dL. -->


heart_disease$target <- ifelse(heart_disease$num == 0, 0, 1)

t.test(thalch ~ target, data = heart_disease)
# <!-- p is low → null will go: Null hypothesis is rejected. -->
# <!-- Sample means are significantly different. -->
# <!-- Patients with heart disease have a significantly lower average maximum heart rate compared to those without heart disease. -->

table(heart_disease$cp)
#Anova:  chol ~ as.factor(cp)

anova_result <- aov(chol ~ as.factor(cp), data = heart_disease)
summary(anova_result)
# <!-- Since the p-value (0.0000598) is less than the typical alpha level of 0.05 (or even 0.01 or 0.001), we reject the null hypothesis. -->
# <!-- there's a significant difference in cholesterol across chest pain types. -->

#Simple Linear Regression
regression_model = lm(chol ~ age, data = heart_disease)
summary(regression_model)
# <!-- The coefficient for age is -1.017, meaning for every 1-year increase in age, the cholesterol decreases by ~1.02 units, on average. -->
# <!-- The p-value for age = 0.0101 -->
# <!-- p is low → null will go: Null hypothesis is rejected. -->
# <!-- This means the effect of age on cholesterol is statistically significant. -->
# <!-- R-squared = 0.0074 -->
# <!-- Only 0.74% of the variability in cholesterol is explained by age — the model has a weak predictive power. -->
# <!-- F-statistic = 6.653, p = 0.01006 -->

table_cp_target <- table(heart_disease$cp, heart_disease$target)
chi_result <- chisq.test(table_cp_target)
chi_result
# <!-- Tested whether there is a significant association between chest pain type (cp) and heart disease presence (target). -->
# <!-- X² = 102.56 -->
# <!-- df = 3 -->
# <!-- p-value < 2.2e-16 -->
# <!-- Interpretation: -->
# <!-- p is low → null will go: Null hypothesis is rejected. -->
# <!-- There is a significant association between chest pain type and heart disease. -->






